### 2023-10-15 scratch/gm_addition_to_cv-edited_v2.Rmd
Further edits to mixed model portion of vignette

---
title: "Cross-validation of regression models: mixed model examples (edited v2)"
author: "John Fox and Georges Monette"
date: "`r Sys.Date()`"
package: cv
output:
  rmarkdown::html_vignette:
  fig_caption: yes
bibliography: ["cv.bib"]
csl: apa.csl
vignette: >
  %\VignetteIndexEntry{Cross-validation of regression models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.align = "center",
  fig.height = 6,
  fig.width = 7,
  fig.path = "fig/",
  dev = "png",
  comment = "#>" #,
  # eval = nzchar(Sys.getenv("REBUILD_VIGNETTES"))
)

# save some typing
knitr::set_alias(w = "fig.width",
                 h = "fig.height",
                 cap = "fig.cap")

# colorize text: use inline as `r colorize(text, color)`
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}

.opts <- options(digits = 5)

knitr::opts_chunk$set(comment="    ")
```

```{r packages, include=FALSE}
library(cv)
library(car)
library(lattice)
library(latticeExtra)
library(lme4)
```
The following artificial data set exemplifies striking aspects of cross-validation
with hierarchical models. Model comparisons using cluster-based versus individual-based 
cross-validation may not agree on a best model. Furthermore, commonly-used measures
of fit, such as least-squares, do not necessarily become smaller as 
models becomes larger.

Varying the parameters of the model, particularly the between-subject variance
relative to the within-subject variance, can also reveal strong differences in the
behaviour of different algorithms used to fit mixed models. These
issues are not pursued in any depth in this vignette.

This example exhibits both phenomena described above. Consider a researcher studying the improvement
in a skill, yodelling for example, among students enrolled in a four-year
yodelling program. The plan is to measure each student's skill level
at the beginning of the program and every year thereafter until
the end of the program, resulting in five annual measurements for each student.

It turns out that yodelling appeals to students of all ages.
Students enrolling in the program range in age from 20 to 70.
Moreover, participants' untrained yodelling skill is similar at all ages, as is
their rate of progress with training. All students complete the
five-year program.

The researcher, with more expertise in yodelling than modelling,
decides to model the response $Y$, yodelling skill, against
age, $X$, reasoning that students get older during their stay in the program,
and age can serve as a proxy for elapsed time.
The researcher knows that a mixed model should be used to account for clustering
due to the expected similarity of measurements taken from each student.

The figures below show the pattern of responses expected from a representative
subset of individuals.

The between-student effect of age is zero but the within-student effect is 1.
Due to the large variation in ages between students, a least-squares model
fitting age as a predictor would produce an estimated slope that is close to
zero, since the estimated slope is heavily weighted toward the between-student effect.
Consider a mixed model using the same fixed-effects model with the addition of a random intercept.

Depending on the estimated variance parameters of the model, a mixed model will apply 
varying degrees of
shrinkage to the random intercept BLUPs [insert reference to ARA&GLM] that
correspond to the variation in the heights of the parallel fitted lines for each student, measured
at a common value of age.

In the example below, the model applies little shrinkage, allowing a large variability in the heights
of fitted lines, which approach closely the observed values for each student.
Consequently, the model's fit is similar to that of
a fixed-effects model with an effect of age together with a categorical predictor for 
individual students. Thus, the model fits individual observations well, resulting in a
favourable assessment using individual-based cross-validation. In contrast, the very large
variability in the BLUPs results in large residuals for
cluster-based cross-validation, producing an
unfavourable assessment compared with the smaller model using only a random intercept.

By comparison, if the mixed model had applied considerable shrinkage, neither the
between-cluster nor the between-individual cross-validations would have shown much 
improvement over the random-intercept only model.  In our experience, the degree of
shrinkage does not vary smoothly as parameters are changed but tends to be
"all or nothing." Its behaviour is affected considerably by the choice of the
algorithm used to fit the model.

Finally, a model that also includes a compositional variable consisting of the mean age
of each student, results in the random intercept BLUPs representing variation in the students' fitted
lines measured at the mean age of each student. Thus, with this example, both 
the cluster-based and the individual-based fits are relatively good.

To serve as a comparison, we also include cross-validation using the model known to 
have generated the data, namely a model with a between-subject age effect equal to zero by
modelling the response $Y$ with a fixed effects model using an age centered within
each student, i.e. the age at assessment less the mean age of each student's 
assessments, plus a random term for each student.

A closely related question is that of non-convergence of mixed-model
algorithms. Adjusting the parameters generating this data can be used to create examples
illustrating problems of non-convergence with hierarchical models.
Fitted models with relatively small between-cluster variance, even those using simple
random intercept models, are prone to non-convergence due to the
nature of optimizing algorithms used in `lme` and `lmer`.

In some contexts, non-convergence,
can be viewed as a diagnostic of a poor model. However, with a simple
random intercept model where it may indicate a small between-cluster
variance, non-convergence is an annoyance because it would be statistically correct to
report that the data are consistent with a zero between-cluster population variance,
i.e. that observed between-cluster variance in the data can be explained by
the between-individual variability within clusters.  This leads to
valid statistical inferences. Although the calculation of accurate p-values for hypotheses concerning
random effects may be complicated
by the fact that the estimated variance is at the boundary of the parameter space,
the estimated variance-covariance matrix of fixed-effect parameters will generally
be non-singular and may be used for valid Wald tests of linear hypotheses
of fixed-effect parameters.

Using cross-validation based on fitting random subsets compounds the problem
since subsets for which there is non-convergence cause the cross-validation to
fail unless non-converging subsets are omitted, which may bias results since non-convergence
is likely to be related to the estimated between-cluster variances.

Varying the parameter for between-cluster variance (`SDre` in the code below),
will generate data sets with varying propensities to produce
non-convergence. These data sets can be used to explore the consequences of non-convergence and to
compare the performance of different optimizers.

At present, the results below are based on a data set that does not
produce non-convergence and are limited to comparisons obtained
by varying two factors:

- Four models: random intercept only, adding x as a predictor, adding the 
  group means of x (xm) as a predictor, and
  a fourth model with a random intercept and x centered within each student.
- Cross-validation resampling individual cases or resampling clusters.

Since the independence of clusters may be a reasonable assumption, as is the
the independence of individual observations from the same cluster *conditional*
on clusters, each form of resampling addresses the
suitability of models
to answer different questions.
Results based on individual resampling generalize to new samples from the same
clusters. Resampling clusters generalizes to new samples of clusters and
cases.

We start by generating and plotting the data.

```{r data}
# Generate data set

# Parameters:
set.seed(9693) 
Nb <- 100    # number of groups
Nw <- 5      # number of individuals within groups
Bb <- 0      # between-group regression coefficient on group mean
SDre <- 2.0  # between-group SD of random level relative to group mean of x
SDwithin <- 0.5  # within group SD
Bw <- 1     # within group effect of x
Ay <- 10    # intercept for response
Ax <- 20    # starting level of x
Nx <- Nw*10 # number of distinct x values

Data <- data.frame(
  group = factor(rep(1:Nb, each=Nw)),
  x = Ax + rep(1:Nx, length.out = Nw*Nb)
) |>
  within(
    {
      xm  <- ave(x, group, FUN = mean) # within-group mean
      y <- Ay +
        Bb * xm +                  # contextual effect
        Bw * (x - xm) +            # within-group effect
        rnorm(Nb, sd=SDre)[group] +  # random level by group
        rnorm(Nb*Nw, sd=SDwithin)    # random error within groups
    }
  )
```

```{r plot1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Hierarchical data set, showing first 10 of 100 clusters."
plot <- xyplot(y ~ x, data=Data[1:Nx, ], group=group,
               ylim=c(4, 16),
               par.settings=list(superpose.symbol=list(pch=1, cex=0.7))) +
    layer(panel.ellipse(..., center.cex=0))
plot
```

The Y variable could represent scores on a programming aptitude test of 10 students
of varying ages, each tested
annually during a 5-year training program.  The X variable is the age
at which each test occurred.

We next obtain six fits with three models and two fitting functions.
```{r}
# random intercept only:
fit0 <- lmer(y ~ 1 + (1 | group), Data)

# effect of x and random intercept:
fit1 <- lmer(y ~ x + (1 | group), Data)

# effect of x, mean of x, and random intercept:
fit2 <- lmer(y ~ x + xm + (1 | group), Data)
        # equivalent to y ~ I(x - xm) + xm + (1 | group)

# model generating the data (where Bb = 0)
fit3 <- lmer(y ~ I(x - xm) + (1 | group), Data)
```
We can compare estimated fixed-effect coefficients.

```{r, message=FALSE,warning=FALSE}
compareCoefs(fit0,  fit1, fit2, fit3)
```

We visualize predicted values using population predictions with the fixed effects model only
and using BLUPs for random effects. 
```{r}
Data <- within(Data, {
  fit_mod0.fe <- predict(fit0, re.form = ~ 0)
  fit_mod0.re <- predict(fit0)
  fit_mod1.fe <- predict(fit1, re.form = ~ 0)
  fit_mod1.re <- predict(fit1)
  fit_mod2.fe <- predict(fit2, re.form = ~ 0)
  fit_mod2.re <- predict(fit2)
  fit_mod3.fe <- predict(fit3, re.form = ~ 0)
  fit_mod3.re <- predict(fit3)
})

# prepare for plotting:
Data_long <- reshape(Data[1:Nx, ], direction = "long", sep = ".", 
              timevar = "effect", varying = grep("\\.", names(Data[1:Nx, ])))
Data_long$id <- 1:nrow(Data_long)
Data_long <- reshape(Data_long, direction = "long", sep = "_", 
              timevar = "modelcode",  varying = grep("_", names(Data_long)))
Data_long$model <- factor(
  c("~ 1", "~ 1 + x", "~ 1 + x + xm", "~ 1 + I(x - xm)")
  [match(Data_long$modelcode, c("mod0", "mod1", "mod2", "mod3"))]
)
```

```{r plot-fits-mod0}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values from random intercept model."
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" & effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod0" &  effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  main="Model: y ~ 1 + (1 | group)",
  key=list(
    corner=c(0.05, 0.05),
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod1}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values from model with random intercept and x as a predictor."
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod1" & effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  main="Model: y ~ 1 + x + (1 | group)",
  ylim=c(-15, 35),
  key=list(
    corner=c(0.95, 0.05),
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod2}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values from model with random intercept, x, and the group mean of x as predictors."
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" & effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod2" & effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  main="Model: y ~ 1 + x + xm + (1 | group)",
  ylim=c(4, 16),
  key=list(
    corner=c(0.05, 0.05),
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```
```{r plot-fits-mod3}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Fitted values from the estimated model generating the data."
(plot +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod3" & effect == "fe"),
         groups=group, type="l", lwd=2) +
  xyplot(fit ~ x, subset(Data_long, modelcode == "mod3" & effect == "re"),
         groups=group, type="l", lwd=2, lty=3)
) |> update(
  main="Model: y ~ 1 + I(x - xm) + (1 | group)",
  ylim=c(4, 16),
  key=list(
    corner=c(0.05, 0.05),
    text=list(c("fixed effects only","fixed and random")),
    lines=list(lty=c(1, 3))))
```

### Cross-validation of models

We carry out cross-validation resampling individuals within clusters, based on
fixed and predicted random effects, and resampling clusters, based on
fixed effects only.
In order to reduce between-model variability in comparisons of models,
the `cv` method applied to the list of models created by the `models` function,
performs cross-validation with the same training and validation sets
for each model.

```{r cross-validation-clusters}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Cross-validation comparing random intercept models with varying fixed effects and with cluster-based resampling."
modlist <- models("~ 1"=fit0, "~ 1 + x"=fit1, 
                  "~ 1 + x + xm"=fit2, "~ 1 + I(x - xm)"=fit3)
cvs_clusters <- cv(modlist, data=Data, cluster="group", k=10, seed=6449)
plot(cvs_clusters, main="Model Comparison, Cluster-Based CV")
```
```{r cross-validation-cases}
#| out.width = "100%",
#| fig.height = 6,
#| fig.cap = "Cross-validation comparing random intercept models with varying fixed effects and with case-based resampling."
cvs_cases <- cv(modlist, data=Data, seed=9693)
plot(cvs_cases, main="Model Comparison, Case-Based CV")
```

In summary, the model with $X$ alone, without the compositional mean of $X$, is assessed as
fitting very poorly sampling clusters but, relatively, much better resampling individual cases.
The model that includes both $X$ and its compositional mean produces better results using both
criteria: resampling individual cases and resampling cluster, illustrating the need to
assess models at different levels of resampling.