### 2023-10-29 scratch/notes_john/simple-var-selection.Rmd
restore scratch dir with notes_john subdir; update tarball under scratch

Now suppose that we use a simple-minded variable-selection procedure to pick the $p' = 5$ predictors that are most highly correlated with $y$ and use them in a least-squares regression:
```{r pick-best-5}
r <- as.vector(cor(X, y))
names(r) <- 1:100
(best5 <- sort(r, decreasing=TRUE)[1:5])
m <- lm(y ~ X[, names(best5)])
summary(m)
mse(y, fitted(m))
```
The $R^2$ for the selected model is small, but the omnibus $F$-test that all of the 5 $\beta$s are 0 is (spuriously) "statistically significant," as are the regression coefficients for 2 of the 5 selected $x$s. These results are, of course, nonsense. Notice that the mean-squared error of prediction for the model is $0.962$, which is smaller than the true error variance $\sigma^2 = 1$. The estimated error variance for the least-squares regression is also too small, $\widehat{\sigma}^2 = 0.984^2 = 0.968$.

Now let us cross-validate the MSE for the selected model:
```{r cv-best-5}
cv(m, seed=4354)
```
Cross-validation consequently also over-states the predictive power of the model because, to reiterate, the model was selected on the same data used in the CV process.

How then can one proceed correctly to cross-validate a model-selection procedure? The model-selection procedure should be applied with each fold removed, and the resulting selected model applied to predict the response for the cases in the omitted fold.

For our simple variable-selection procedure this process can be automated:
```{r cvSelect}
selectBest <- function(data, n.best, y, x){
  # data: data frame containing all of the data
  # n.best: number of xs to select
  # y: name or index of the response
  # x: vector of names or indices of candidate predictors
  y <- data[ , y]
  X <- as.matrix(data[ , x])
  colnames(X) <- 1:ncol(X)
  r <- as.vector(cor(X, y))
  names(r) <- 1:ncol(X)
  (best <- sort(r, decreasing=TRUE)[1:n.best])
  best <- names(best)
  X <- X[, best]
  data <- data.frame(y, X)
  lm(y ~ X, data=data)
  }

D <- data.frame(y, X)
selectBest(data=D, n.best=5, y=1, x=2:101)
```
Next, we write a function to apply `selectBest()` with a fold removed:
```{r cvSelectBest}
cvSelectBest <- function(data, indices, n.best, y, x){
  # data: data frame containing all of the data
  # indices: row indices in data for the current fold
  #          if omitted, the function returns the MSE
  #          for the full data set
  # n.best: number of xs to select
  # y: name or index of the response
  # x: vector of names or indices of candidate predictors
  if (missing(indices)) {
    model.i <- selectBest(data, n.best=n.best, y=y, x=x)
    fit.o.i <- fitted(model.i)
    return(mse(data[, y], fit.o.i))
  }
  data.minus.fold <- data[-indices, ]
  model.i <- selectBest(data.minus.fold, n.best, y, x)
  X <- cbind(1, as.matrix(data[, names(coef(model.i))[-1]]))
  fit.o.i <- as.vector(X %*% coef(model.i))
  fit.i <- fit.o.i[indices]
  c(mse(data[indices, y], fit.i), mse(data[, y], fit.o.i))
}
```
For example:
```{r try-cvSelectBest}
cvSelectBest(data=D, n.best=5, y=1, x=2:101) # whole data set
cvSelectBest(data=D, indices=1:100,
             n.best=5, y=1, x=2:101) # omit first 100 cases
```

We wrote `cvSelectBest()` so that it's useable  with the `cvSelect()` function in the **cv** package. `cvSelect()` applies a model-selection procedure, such as our `cvSelectBest()`, in $k$-fold or LOO CV.

`cvSelectBest()` has two required arguments: `procedure`, a model-selection function whose first two arguments should be `data`, a full data frame, and `indices`, the case indices for the current fold. The user can also supply `k` and `seed` arguments to `cvSelectBest()`; as for `cv()`, `k` defaults to `10`, and if `seed` isn't supplied, a random-number generator seed is randomly selected and saved. Any other arguments to `cvSelect()` are passed through to the model-selection function.

So, we can cross-validate our simple model-selection procedure applied to the randomly generated data by
```{r cv-select-best-5}
cvSelect(cvSelectBest, data=D, seed=321,
         n.best=5, y=1, x=2:101)
```
Properly cross-validating model selection for the example produces a realistic estimate of the poor predictive power of the selected model, with the estimated MSE of prediction very slightly larger than the marginal variance of $y$, that is, $\sigma^2 = 1$.