### 2024-09-22 scratch/JSS-submission/Revision-1/article-add.Rmd
additions re Python

---
documentclass: jss
author:
  - name: John Fox
    orcid: 0000-0002-1196-8012
    affiliation: 'McMaster University'
    address: |
      | Hamilton, Ontario, Canada
    email: \email{jfox@mcmaster.ca}
    url: https://www.john-fox.ca/
  - name: Georges Monette
    orcid: 0000-0003-0076-5532
    affiliation: 'York University'
    address: |
      | Toronto, Ontario, Canada
    email: \email
    url: http://blackwell.math.yorku.ca/gmonette
title: DRAFT Addition re Other Packages
preamble: >
  \usepackage{amsmath}
output: 
  rticles::jss_article:
    extra_dependencies: "subfig"
bibliography: ["cv-add.bib"]
---

# Comparing cv to other software for cross-validation

We have examined provisions for cross-validation
in \proglang{SAS}, \proglang{Stata}, and \proglang{Python} although
we do not have recent experience with languages other than \proglang{R}.

In \proglang{Python} [@pedregosaScikitlearnMachineLearning:2011], 
statistical modelling is supported through extensive libraries
such as \pkg{scikit-learn} [@ScikitLearnDevelopers:2024]
and \pkg{Tensorflow} [@tensorflow:2024].  
Functions for cross-validation are provided in the \pkg{Scikit-learn}
library that includes functions
for k-fold, repeated k-fold, leave-one-out, leave-p-out cross-validation.
These functions can be applied to data sets consisting of independent observations,
hierarchical groups, or time series data. Documentation
for the \pkg{scikit-learn} provides
a sample program [@ScikitLearnDevelopers:2024] to show
users how to implement nested cross-validation.